---
title: "House Prices"
author: "Asger Rasmussen"
format: pdf
editor: visual
---

```{r}
#| include: false

knitr::opts_chunk$set(echo = TRUE)
library(fable)
library(tidyverse)
library(readxl)
library(janitor)
library(lubridate)
library(zoo)
library(ggthemes)
library(tsibble)
library(feasts)
library(viridis)
theme_set(theme_fivethirtyeight())


```

# Introduction

This paper seeks to forecast house prices, on the communal level of Denmark. I will run multiple forecasting models, and compare which ones seems to give me the best result.

## Theory

When you click the **Render** button a document will be generated that includes both content and the output of embedded code. You can embed code like this:

## Data

House prices, are quarterly detached/terrace house prices on Region level, the data spans from 1992 until 2023, the realized transaction priced is used.

Pre tax Income for people (14 years +), by Region, yearly data from 1987 until 2022 for both men and women. The unit is Average income for persons in the group (DKK).

Population by region

```{r - BM010: Property prices in housing market by area (1992Q1-2023Q3) Kommune}
#| echo: false

House_Prices_raw <- read_excel("Data/House Price Region Quarter.xlsx")

House_Prices <- House_Prices_raw %>% 
  filter(!row_number() %in% c(1)) %>% 
  row_to_names(., 1) %>% 
  clean_names() %>% 
  as_tibble() %>% 
  rename(Date = na,
         Category = na_2,
         Type = na_3,
         Capital = capital_region,
         Zealand = region_zealand,
         "Central Denmark" = central_denmark_region,
         "Sourth Denmark" = region_of_southern_denmark,
         "North Jutland" = region_of_north_jutland
         ) %>% 
  mutate(Quarter = yearquarter(Date)) %>% 
  select(-Category, -Type, -Date) %>% 
  pivot_longer(
    cols = Capital:"North Jutland",
    names_to = "Region",
    values_to = "Price"
  ) %>% 
  mutate(Price = as.integer(Price)) %>% 
  as_tsibble(key = Region, index = Quarter)

Average_income_raw <- read_excel("Data/Average Pre-Tax Income Region Yearly.xlsx")

Pre_tax_income <- Average_income_raw |> 
  filter(!row_number() %in% c(1,2)) |>
  select(-...2, -...3, -...4) |> 
  rename(
    Date = "Income, total by time, income interval, sex, unit and region",
    "Region Hovedstaden" = ...5,
    "Region Sjælland" = ...6,
    "Region Syddanmark" = ...7,
    "Region Midtjylland" = ...8,
    "Region Nordjylland" = ...9
  ) |> 
  mutate(Date = as.Date(paste0(Date, "-01-01"))) |> 
  pivot_longer(
    cols = "Region Hovedstaden":"Region Nordjylland",
    names_to = "Region",
    values_to = "Income"
  ) |>
  mutate(Year = yearmonth(Date)) |> 
  filter(!is.na(Income)) |> 
  select(-Date) |> 
  as_tsibble(key = Region, index = Year)

Population_raw <- read_excel("Data/Population Region Quarter.xlsx")

Population <- Population_raw |> 
  filter(!row_number() %in% c(1,2)) |> 
  select(-...2, -...3, -...4) |> 
   rename(
    Date = "Population at the first day of the quarter by time, marital status, age, sex and region",
    "Region Hovedstaden" = ...5,
    "Region Sjælland" = ...6,
    "Region Syddanmark" = ...7,
    "Region Midtjylland" = ...8,
    "Region Nordjylland" = ...9
  ) |> 
  pivot_longer(
    cols = "Region Hovedstaden":"Region Nordjylland",
    names_to = "Region",
    values_to = "Population"
  ) |> 
  mutate(Quarter = yearquarter(Date),
         Population = as.integer(Population)) |> 
  select(-Date) |> 
  as_tsibble(key = Region, index = Quarter)

Construction_raw <- read_excel("Data/Construction Permits Region Quarter.xlsx")

Permits <- Construction_raw |> 
  filter(!row_number() %in% c(1,2)) |> 
  select(-...2, -...3, -...4) |> 
  rename(
    Date = "Residential Construction (not adjusted for delays) by time, use, phase of construction, builders and region",
    Region_Hovedstaden = ...5,
    Region_Sjælland = ...6,
    Region_Syddanmark = ...7,
    Region_Midtjylland = ...8,
    Region_Nordjylland = ...9
  ) |> 
  fill(Date) |> 
  mutate(Region_Hovedstaden = as.integer(Region_Hovedstaden),
         Region_Midtjylland = as.integer(Region_Midtjylland),
         Region_Sjælland = as.integer(Region_Sjælland),
         Region_Syddanmark = as.integer(Region_Syddanmark),
         Region_Nordjylland = as.integer(Region_Nordjylland)) |> 
  group_by(Date) |>
  summarise(across(starts_with("Region_"), sum, na.rm = TRUE)) |> 
  pivot_longer(
    cols = Region_Hovedstaden:Region_Nordjylland,
    names_to = "Region",
    values_to = "Permits"
  ) |> 
  ungroup() |> 
  filter(!row_number() %in% c(361:365)) |> 
  mutate(Quarter = yearquarter(Date)) |> 
  select(-Date) |> 
  as_tsibble(key = Region, index = Quarter)

Houses_Sold_raw <- read_excel("Data/Dwellings sold Region Quarter.xlsx")

Houses_Sold <- Houses_Sold_raw |> 
  filter(!row_number() %in% c(1,2)) |> 
  select(-...2, -...3) |> 
  rename(
    Quarter = "Movements in housing market by time, type of movement, property category and area",
    Region_Hovedstaden = ...4,
    Region_Sjælland = ...5,
    Region_Syddanmark = ...6,
    Region_Midtjylland = ...7,
    Region_Nordjylland = ...8
  ) |> 
  pivot_longer(
    cols = Region_Hovedstaden:Region_Nordjylland,
    names_to = "Region",
    values_to = "Houses_Sold"
  ) |> 
  mutate(Houses_Sold = as.integer(Houses_Sold),
         Quarter = yearquarter(Quarter)) |> 
  as_tsibble(key = Region, index = Quarter)

Houses_for_sale_raw <- read_excel("Data/Houses for Sale Region Quarter.xlsx")

Houses_for_sale <- Houses_for_sale_raw |> 
  filter(!row_number() %in% c(1,2)) |> 
  select(-...2, -...3) |> 
  rename(
    Quarter = "Movements in housing market by time, type of movement, property category and area",
    Region_Hovedstaden = ...4,
    Region_Sjælland = ...5,
    Region_Syddanmark = ...6,
    Region_Midtjylland = ...7,
    Region_Nordjylland = ...8
  ) |> 
  pivot_longer(
    cols = Region_Hovedstaden:Region_Nordjylland,
    names_to = "Region",
    values_to = "Houses_for_sale"
  ) |> 
  mutate(Houses_for_sale = as.integer(Houses_for_sale),
         Quarter = yearquarter(Quarter)) |> 
  as_tsibble(key = Region, index = Quarter)
```

## Univariate analysis

Let us look at house prices, and see how well we can forecast using only a single time series.

```{r}

House_Prices |> 
  filter(Region == "Capital") |> 
  ggplot(aes(x = Quarter, y = log(Price))) +
  geom_line(color = "#31A354", size = 1.3) +
  theme_fivethirtyeight() +
  labs(
    title = "Capital House Prices by Quarter 1992 - 2023"
  )

```

From this graph it is clear to see we have a positive trend within our data, there also seems to be some cyclic behaviour. The housing bubble and Covid - 19? Let us take a closer look that the seasonality

```{r}

House_Prices |> 
  filter(Region == "Capital") |>
  gg_season(log(Price))

House_Prices |> 
  filter(Region == "Capital") |> 
  gg_subseries(log(Price))

```

There seems to be no strong signs of seasonality within our data for house prices. Let us take a look at the lags plot for this single time series

```{r}

House_Prices |> 
  filter(Region == "Capital") |> 
  gg_lag(Price, geom = "point") +
  theme_minimal() +
  scale_color_viridis(discrete = TRUE, option = "viridis") +
  labs(x = "lag(House, k)")

```

Let us take a look at autocorrelation

```{r}


House_Prices |>
  filter(Region == "Capital") |> 
  ACF(log(Price), lag_max = 36) |> 
  autoplot() +
  labs(
    title = "ACF House Price in the Capital Region"
  )

House_Prices |> 
  filter(Region == "Capital") |> 
  PACF(log(Price), lag_max = 36) |> 
  autoplot() +
  labs(
    title = "PACF House Price in the Capital Region"
  )
```

It seems to die out around 32 lags out, which is about 8 years so there is a high persistence in our Auto regressive of the time series. The PACF tells a different story that only the first lag is significant, this leads us that an AR(1) seems to fit this data best.

```{r}



```
